---
title: "2023-08-29 qPCR Analysis"
author: "Dan Rice"
format:
  html:
    code-fold: false
---
## Objectives

- Compare several wastewater filtering options by measuring the nucleic acid content by qPCR. [Experimental design here](https://docs.google.com/document/d/10aEcOwlJ0ZZANC93gd5KM6j-u8T8MI4ML7Ibh5uivMo/edit?usp=sharing).
- Get Dan up to speed with working with this data.
- Explore options for data analysis workflows.

## Preliminary work

- Ari put the `.eds` files into Google Drive from the lab computer. (He also exported some Excel files but we're not using those.)
- Dan installed the Google Drive desktop app and Design and Analysis on his Mac, opened the `.eds` files, and fixed some missing data in the plate layout
- Dan used the "Analyze" in Design and Analysis to automatically calculate thresholds and compute `c_q` values.
- Dan exported the data to `.csv`.
- Symlinked the google drive folder for the airport experiment to `~/airport/` on his computer so I don't have to refer to the whole filepath here.

## Data import

```{r}
library(readr)
library(dplyr)
library(purrr)

data_dir <- "~/airport/[2023-08-29] CP Prefilters vs Vacuum Filters/Test 1 qPCR Results/csv/"
filename_pattern <- "Results"
```

```{r}
data_raw <- list.files(
                       data_dir,
                       pattern = filename_pattern,
                       full.names = TRUE
                       ) |>
  map(function(f) read_csv(f, skip=23)) |>
  list_rbind()
print(data_raw)
```

```{r}
tidy_data <- data_raw |>
  mutate(Cq = as.double(Cq))
print(tidy_data$Target)
```

```{r}
library(ggplot2)
tidy_data |>
  ggplot(mapping = aes(x=Cq, y=Sample)) +
  geom_point() +
  facet_wrap(facets = ~Target)
```

## TODO

- Figure out what the `Sample` numbers mean with respect to the different treatments
- There are a few NaNs that show up as missing points and aren't evident in the plot. Will investigate later.
