---
title: "2023-10-05 Qubit data exploration"
author: "Dan Rice"
date: 2023-10-05
toc: true
format:
  html:
    code-fold: false
---
## Objectives

Trying to understand the Qubit output.

## Preliminary work

Ari exported the data from a recent (somewhat arbitrarily selected) run to a flash drive and put it [here](https://drive.google.com/drive/folders/1Qr_XHXNTNR1z33_XhSoinrYK7o-wdVdT?usp=drive_link).

Olivia looked up the concentration of the standards and [said](https://twist.com/a/197793/ch/591292/t/5608437/c/86173146):

> The concentration of Std 1 is 0 ng/uL (TE buffer). Std 2 is 10ng/uL of rRNA (also in TE buffer).

The Assay Kit User guide is [here](https://assets.fishersci.com/TFS-Assets/LSG/manuals/Qubit_RNA_HS_Assay_UG.pdf)

## Data import

```{r}
library(here)
library(tidyverse)
data_dir <- here("_data", "qubit")
data <- read_csv(here(data_dir, "QubitData_10-05-2023_08-32-08.csv"))
glimpse(data)
```

## Concentration curves

My first guess is that the relationship between concentration and RFU should be linear.

```{r}
data |>
  ggplot(mapping = aes(x = `Far Red RFU`, y = `Qubit tube conc.`)) +
  geom_point() +
  geom_smooth(method = "lm")
```

It is not.

```{r}
fit <- lm(`Qubit tube conc.` ~ `Far Red RFU`, data)
print(fit)
print(fit$residuals)
```

To be continued...
