---
title: "2023-07-18 Extraction Experiment 1 qPCR Analysis"
author: "Dan Rice"
date: 2023-09-20
toc: true
format:
  html:
    code-fold: false
---
## Objectives

Testing the efficacy of RNA extraction kits in influent samples.
See [experiment google doc](https://docs.google.com/document/d/1KOAolQzHvCE8dHVn9bxrBJwm7sLAHeJQQjzc_lslBf0/edit?usp=sharing).

## Preliminary work

- Someone else already exported qPCR results as CSV
- I copied the file `meta_samples.csv` and manually added elution volumes
  that I found in [this doc](https://docs.google.com/document/d/1AJfbucGj1frkfl2jfVNSlLgP2pjnAYZwiYy91Oy_wzg/edit#bookmark=id.atrajjxwxf20)
  and saved the results in `meta_samples_with_elution_volumes.csv`.

## Data import

```{r}
library(readr)
library(dplyr)
library(purrr)
library(stringr)
library(ggplot2)
library(tidyr)
library(broom)
```

```{r}
data_dir <-
  "~/airport/[2023-07-18] Extraction-kit comparison 1: Influent/qPCR results/"
filename_pattern <- "_Results_"
```

Not sure why I had to change `skip` from 23 to 22.

```{r}
col_types <- list(
  Target = col_character(),
  Cq = col_double()
)
raw_data <- list.files(
  data_dir,
  pattern = filename_pattern,
  recursive = TRUE,
  full.names = TRUE,
) |>
  print() |>
  map(function(f) {
    read_csv(f,
      skip = 22,
      col_types = col_types,
    )
  }) |>
  list_rbind()
print(raw_data)
```

```{r}
raw_data |> count(Target)
```

```{r}
metadata_file <- paste0(data_dir, "meta_samples_with_elution_volumes.csv")
metadata <- read_csv(metadata_file)
glimpse(metadata)
```

```{r}
tidy_data <- raw_data |>
  mutate(
    replicate = str_extract(Sample, "[A-Z]$"),
  ) |>
  left_join(
    metadata,
    by = join_by(Sample == sample_qPCR)
  )
glimpse(tidy_data)
```

```{r}
tidy_data |>
  count(Sample, Kit, LPA, Target, replicate) |>
  print(n = Inf)
```

## Kit comparison

```{r}
tidy_data |>
  filter(!is.na(Kit)) |>
  ggplot(mapping = aes(
    x = Cq,
    y = Kit,
    color = interaction(replicate, LPA),
  )) +
  stat_summary(
    fun.min = min,
    fun.max = max,
    fun = median,
    position = position_dodge(width = 0.2),
    size = 0.2
  ) +
  facet_wrap(facets = ~Target, scales = "free_x")
```

### Adjusting for elution volume

Assume the amplification efficiency is 100%, so that
an increase in initial concentration by a factor of $x$
decreases $C_q$ by $\log_{2}(x)$.

If a method has elution volume $v$ and we dilute it to total volume $V$,
this reduces its final concentration by a factor $v / V$.
We can put different
methods on the same footing by adding $\log_{2}(v/V)$ to $C_q$
(so that large elution volumes are penalized with a higher adjusted $C_q$).

```{r}
final_volume <- 100
tidy_data |>
  mutate(elution_adjusted_Cq = Cq + log2(Elution_volume / final_volume)) |>
  filter(!is.na(Kit)) |>
  ggplot(mapping = aes(
    x = elution_adjusted_Cq,
    y = Kit,
    color = interaction(replicate, LPA),
  )) +
  stat_summary(
    fun.min = min,
    fun.max = max,
    fun = median,
    position = position_dodge(width = 0.2),
    size = 0.2
  ) +
  facet_wrap(facets = ~Target, scales = "free_x")
```
